vegan_analysis

### 2023-06-01 scripts/4_vegan.qmd
secondary sampling scheme, revised manuscript for ncc

```{r packages}
library("vegan")
```

@@

if (!file.exists(bec_sample_loc)) {
  bec_samples_e <- extract(all_rasts, bec_samples, bind = T)
  writeVector(bec_samples_e, bec_sample_loc, overwrite = T)
}

@@

bec_normal_ndup <- bec_df %>%
  group_by(across(c(bec, category, basal_area:total_biomass))) %>%
  dplyr::summarize(across(ends_with("DHI"), mean)) %>%
  ungroup() %>%
  mutate(across(basal_area:VarDHI,  .fns = ~ normal(.x))) 
```

```{r global-model}
to_rda <- bec_normal %>%
  select(-elev_mean,
         -percentage_first_returns_above_mean,
         -elev_stddev)

@@

modelled <- to_rda %>%
  select(loreys_height,
         basal_area,
         total_biomass,
         gross_stem_volume)

rda_out <- 
  rda(dhi_tab ~ elev_p95 + percentage_first_returns_above_2m + elev_cv + # simple/raw/extracted
      basal_area + gross_stem_volume + loreys_height + total_biomass, # modelled
      data = to_rda)

rda_out

@@

rda_eig <- rda_out$CCA$eig

@@

summary(rda_out)$sites %>% 
  as.data.frame() %>% 
  select(RDA1) %>% 
  bind_cols(to_rda) %>%
  select(-bec, -category) %>% 
  cor() %>%
  as.data.frame() %>%
  select(RDA1) %>%
  arrange(desc(abs(RDA1)))

@@

```{r bec-nest}
to_nest <- bec_normal

bec_nested <- to_nest %>%
  group_by(bec) %>%
  nest() %>%
  ungroup()
```

@@

  rda(dhi_tab ~ elev_p95 + percentage_first_returns_above_2m + elev_cv + # simple/raw/extracted
      basal_area + gross_stem_volume + loreys_height + total_biomass, # modelled, put condition here for a conditional rda
      data = tibble)
}

@@

  modelled <- tibble %>%
    select(loreys_height,
           basal_area,
           total_biomass,
           gross_stem_volume)

@@

fracts_out <- outs %>%
  mutate(indfracts = map(var, .f = function(x) {
    x$part$indfract %>%
      mutate(type = c("extracted", "modelled", "overlap", "residual")) %>%
      select(type, Adj.R.squared) %>%
      pivot_wider(names_from = type, values_from = Adj.R.squared)
    }
    )) %>%
  unnest(indfracts) %>%
  arrange(desc(residual)) %>%
  select(bec, extracted, modelled, overlap) %>%
  pivot_longer(c(extracted, modelled, overlap)) %>%
  mutate(value = ifelse(value > 0, value, 0)) %>%
  pivot_wider()
```

```{r raincloud-plot}
library(ggdist)
fracts_out %>%
  left_join(keys$bec, by = c("bec" = "zone")) %>%
  pivot_longer(extracted:overlap) %>%
  mutate(name = str_to_title(name)) %>%
  mutate(value = ifelse(value < 0, 0, value)) %>%
  ggplot(aes(x = value, y = name)) +
  stat_halfeye(
    # adjust bandwidth
    adjust = .5,
    # move to the right
    justification = -0.2,
    # remove the slub interval
    .width = 0,
    point_colour = NA
    ) +
  geom_boxplot(
    width = 0.12,
    # removing outliers
    outlier.color = NA,
    alpha = 0.5
  ) +
  stat_dots(
    # ploting on left side
    side = "left",
    # adjusting position
    justification = 1.1,
    # adjust grouping (binning) of observations
    binwidth = 0.0025,
    #position = "stack",
    aes(fill = group, col = group),
    alpha = 0.5
  ) +

  labs(y = NULL,
       x = "% Variance Explained",
       fill = "BEC Group",
       col = "BEC Group") +
  scale_x_continuous(labels = scales::label_percent()) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.position = c(0.8, .2))

@@

```{r path-setup}
path_base <- tibble(x1 = c(0,
             0,
             0,
             0,
             1),
       y1 = c(1,
              -1,
              1,
              -1,
              0),
       x2 = c(1.85,
              1.85,
              .85,
              .85,
              1.85),
       y2 = c(.15,
              -.15,
              .15,
              -.15,
              0)) %>%
  ggplot(aes(x = x1, y = y1)) +
  geom_segment(aes(xend = x2, yend = y2), arrow = arrow()) +
  geom_text(aes(x = 0, y = -1.15), label = "Modelled") +
  geom_text(aes(x = 0, y = 1.15), label = "Extracted") +
  geom_text(aes(x = .65, y = 0), label = "Overlap") +
  geom_text(aes(x = 2, y = 0), label = "DHIs") +
  geom_point(data = tibble(x1 = c(.85, .85), 
                           y1 = c(-.5, .5)),
             size = 75,
             shape = 21,
             col = "grey") +
  theme_void() +
  scale_x_continuous(limits = c(-0.1, 2.1))

@@

path_base +
  geom_text(data = path_r2_clean, aes(label = r2_label)) +
  labs(title = paste(var_oi, collapse = " "))
```

```{r group-climate-spatial}
climate_folder <- here::here("data", "climate")

clim_vars <- climate_folder %>%
  list.files(pattern = "csv$") %>%
  tools::file_path_sans_ext()

group_clim_avg <-
  here::here(climate_folder, paste0(clim_vars, ".csv")) %>%
  read_csv() %>%
  mutate(var = clim_vars %>% rep(5) %>% sort) %>%
  select(group = BEC_CSV_GR, mean = MEAN, var)

```{r clim-plot-group}
fracts_clims <- fracts_out %>%
  left_join(keys$bec, by = c("bec" = "zone")) %>%
  group_by(group) %>%
  summarize(across(extracted:overlap, .fns = mean)) %>%
  left_join(group_clim_avg, multiple = "all")

clim_var = "tmax"

fracts_clims %>%
  pivot_longer(extracted:overlap) %>%
  filter(var == clim_var) %>%
  ggplot(aes(x = mean, y = value)) +
  geom_smooth(method = "lm", col = "red", lty = "dashed") +
  geom_point() +
  facet_wrap(~str_to_title(name)) +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(x = clim_var,
       y = "% Variance Explained") +
  theme_bw() +
  theme(panel.grid = element_blank())
```

```{r clim-plot-bec}
clim_vars <- climate_folder %>%
  list.files(pattern = "csv$") %>%
  tools::file_path_sans_ext()

bec_clim <- climate_folder %>%
  list.files(pattern = "bec.csv$",
             full.names = T) %>%
  read_csv() %>%
  mutate(var = clim_vars %>% rep(16) %>% sort) %>%
  select(bec = ZONE, mean = MEAN, var)

fracts_out %>%
  pivot_longer(extracted:overlap) %>%
  left_join(bec_clim, multiple = "all") %>%
  filter(var == "ppt") %>%
  ggplot(aes(x = mean, y = value)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  facet_wrap(~name) 
```

```{r testing-idf-live}
to_rda <- bec_normal %>%
  filter(bec == "IDF")

extracted <- to_rda %>%
  select(elev_p95,
         elev_cv,
         percentage_first_returns_above_2m)

modelled <- to_rda %>%
  select(loreys_height,
         basal_area,
         total_biomass,
         gross_stem_volume)

rda_out <- 
  rda(dhi_tab ~ elev_p95 + percentage_first_returns_above_2m + elev_cv + # simple/raw/extracted
      basal_area + gross_stem_volume + loreys_height + total_biomass, # modelled
    data = to_rda)

rda_out

rda_eig <- rda_out$CCA$eig

plot(var_out, Xnames = c("Extracted", "Modelled"))

path_base +
  geom_text(data = path_r2_clean, aes(label = r2_label)) +
  labs(title = "All")

summary(rda_out)$sites %>% 
  as.data.frame() %>% 
  select(RDA1) %>% 
  bind_cols(to_rda) %>%
  select(-bec, -category) %>% cor() %>%
  as.data.frame() %>%
  select(RDA1) %>%
  arrange(desc(abs(RDA1)))
```

@@

```{r cdf-anova}
outs %>% 
  filter(bec == "CDF") %>%
  select(rda) %>% 
  pull(rda) %>%
  .[[1]] %>%
  anova()
```

@@

      anova_out$`Pr(>F)` %>%
        as.data.frame() %>%
        rename(anova_p = 1) %>%
        drop_na() %>%
        mutate(anova_sig = cut(
          anova_p,
          breaks = c(0, 0.001, 0.01, 0.05, .10, 1),
          labels = c("***", "**", "*", ".", "NS")))  %>%
        mutate(axis = paste0("RDA", row_number())) %>%
        relocate(axis)
  }))

anova_sig_axes <- anova_outs %>%
  select(bec, anova) %>%
  unnest(cols = anova) %>%
  filter(anova_sig != "NS")
```

```{r bec-loadings}
out_loadings <- outs %>%
  mutate(loadings = map(rda,.f = function(rda) {
    summary(rda)$sites %>%
      as.data.frame() %>%
      select(starts_with("RDA"))
  }),
  loadings = map2(.x = loadings, .y = data, .f = bind_cols),
  loadings = map(loadings, .f = function(x) {
    rcorr <- x %>% select(-category) %>%
      as.matrix() %>%
      Hmisc::rcorr()

@@

  )) %>%
  select(bec, loadings) %>%
  unnest(loadings)
```

```{r}
out_loadings %>% 
  mutate(axis_sig = cut(
          axis_p,
          breaks = c(0, 0.001, 0.01, 0.05, .10, 1),
          labels = c("***", "**", "*", ".", "NS"),
          right = F)) %>%
  inner_join(anova_sig_axes) %>%
  #filter(axis_sig == "NS") %>%
  arrange(axis_loading) %>%
  count(var)

# need to remove the correlations for variables that arent actually involved
# percentage_first_returns_above_mean, elev_mean, elev_stddev
out_loadings %>% count(var)
```

need to get significant axes from the anova and then filter this

get total rda eigenvalues. find proportion of total for each rda. multiple by 1 - residual. and thats gets path number b/w RD1/RD2 and DHIs. can just filter each by which exist since only a few. goes for path plot

grab the p values w/in zones as well as the correlation coefficients using Hmisc::rcorr(). only plot these plus loadings over threshold

check if deciduous and coniferous have signfiicant axes

>>>>

```{r packages}
library(vegan)
library(tidyverse)
library(cols4all)
library(terra)
library(sf)
library(patchwork)
library(ggradar)
source(here::here("scripts", "get_keys.R"))
```

@@

@@

```{r global-model}
to_rda <- bec_normal %>%
  select(-elev_mean,
         -percentage_first_returns_above_mean,
         -elev_stddev,
         -loreys_height)

forest_types <- bec_normal %>%
  pull(category) %>%
  unique()

# to_rda <- bec_normal %>% 
#   group_by(category) %>% 
#   slice_sample(n = 3000)  %>% 
#   filter(category == "Mixed Wood") %>% 
#   ungroup()

@@

modelled <- to_rda %>%
  select(basal_area,
         total_biomass,
         gross_stem_volume)

rda_out <- 
  rda(dhi_tab ~ elev_p95 + percentage_first_returns_above_2m + elev_cv + # simple/raw/extracted
      basal_area + gross_stem_volume + total_biomass, # modelled
      data = to_rda)

@@

rda_sum <- summary(rda_out)

var_exp <- rda_sum$constr.chi / rda_sum$tot.chi

rda_eig <- rda_out$CCA$eig %>%
  as.data.frame() %>%
  t() %>%
  as.data.frame() %>%
  tibble() %>%
  pivot_longer(cols = everything()) %>%
  mutate(prop = value / sum(value),
         var_exp = prop * var_exp)

@@

rcorr <- summary(rda_out)$sites %>% 
  as.data.frame() %>% 
  select(starts_with("RDA")) %>% 
  bind_cols(to_rda) %>%
  select(-bec, -category) %>% 
  as.matrix() %>%
  Hmisc::rcorr()

all_data_loadings <- left_join(rvals, pvals) %>%
  filter(axis != "RDA3") %>%
  mutate(sig = cut(axis_p, 
                   breaks = c(0, 0.001, 0.01, 0.05, .10, 1),
                   labels = c("***", "**", "*", ".", "NS"),
                   right = F))

for_radar_loads <- all_data_loadings %>%
  filter(!endsWith(var, "DHI")) %>%
  mutate(axis_loading = ifelse(sig == "NS", 0, abs(axis_loading))) %>%
  select(var, axis, axis_loading) %>%
  pivot_wider(names_from = var, values_from = axis_loading) %>%
  mutate(cluster = paste(axis, "- All Data")) %>%
  select(-axis) %>%
  relocate(cluster)

@@

```{r bec-nest}
to_nest <- bec_normal %>%
  select(-percentage_first_returns_above_mean,
         -loreys_height,
         -elev_stddev,
         -elev_mean)

bec_nested <- to_nest %>%
  group_by(bec) %>%
  nest() %>%
  ungroup() %>%
  rename(strata = bec)

# vlce sampling regime to get the samples up
vlce_samples <- read_sf(here::here("data", "samples", "vlce_extracted.shp"))

vlce_df <- vlce_samples %>%
  st_drop_geometry() %>%
  select(-category_1, -category_2) %>%
  rename(category = strata) %>%
  relocate(bec) %>%
  drop_na() %>%
  ungroup()

names(vlce_df) <- names_manual

vlce_normal <- vlce_df %>%
  mutate(across(basal_area:VarDHI,  .fns = ~ normal(.x)))

to_nest <- vlce_normal %>%
  select(-percentage_first_returns_above_mean,
         -loreys_height,
         -elev_stddev,
         -elev_mean)

class_nested <- to_nest %>%
  group_by(category) %>%
  slice_sample(n = 3000) %>%
  nest() %>% 
  ungroup() %>%
  rename(strata = category)

strata_nested <- bind_rows(bec_nested, class_nested)
```

@@

  rda(dhi_tab ~ elev_p95 + percentage_first_returns_above_2m + elev_cv + # simple/raw/extracted
      basal_area + gross_stem_volume + total_biomass, # modelled, put condition here for a conditional rda
      data = tibble)
}

@@

  modelled <- tibble %>%
    select(basal_area,
           total_biomass,
           gross_stem_volume)

@@

fracts_out <- outs %>%
  mutate(indfracts = map(var, .f = function(x) {
    x$part$indfract %>%
      mutate(type = c("extracted", "modelled", "overlap", "residual")) %>%
      select(type, Adj.R.squared) %>%
      pivot_wider(names_from = type, values_from = Adj.R.squared)
    }
    )) %>%
  unnest(indfracts) %>%
  arrange(desc(residual)) %>%
  select(strata, extracted, modelled, overlap) %>%
  pivot_longer(c(extracted, modelled, overlap)) %>%
  mutate(value = ifelse(value > 0, value, 0)) %>%
  pivot_wider()

fracts_out %>% 
  mutate(total = extracted + modelled + overlap) %>%
  arrange(strata %in% forest_types,
          desc(total))
```

```{r}
fract_all <- path_r2 %>% 
  pivot_wider(names_from = type, values_from = Adj.R.squared) %>% 
  select(-residual) %>%
  mutate(strata = "All Data") %>% 
  relocate(strata)

fracts_all <- bind_rows(fracts_out,
                        fract_all)

fracts_tot <- fracts_all %>% 
  mutate(total = extracted + modelled + overlap) %>%
  select(strata, total)

fracts_bar_tab <- fracts_all %>%
  pivot_longer(-strata) %>%
  left_join(keys$bec, by = c("strata" = "zone")) %>%
  mutate(group = ifelse(is.na(group), "Forest Types", group),
         group = ifelse(strata == "All Data", "All", group)) %>%
  left_join(fracts_tot) %>%
  mutate(zone_nm = ifelse(is.na(zone_nm), strata, zone_nm))

fracts_bar_tab

fracts_bar_plot <- fracts_bar_tab %>%
  ggplot(aes(x = value, y = fct_reorder(zone_nm, total), fill = fct_rev(fct_relevel(str_to_title(name), "Overlap", after = 1)))) +
  geom_col(position = "stack") +
  facet_grid(rows = vars(fct_reorder(group, total, .desc = T)),
             scales = "free_y",
             space = "free") +
  labs(x = "Variance Explained",
       y = NULL,
       fill = NULL) +
  theme(legend.position = "bottom",
        panel.spacing.y = unit(0, "lines")) +
  guides(fill = guide_legend(reverse = T)) +
  scale_x_continuous(labels = scales::label_percent())

fracts_bar_plot
ggsave(here::here("outputs", "fracts_bar_plot.png"), plot = fracts_bar_plot)
```

```{r raincloud-plot}
library(ggdist)
fracts_out %>%
  mutate(total = extracted + modelled + overlap) %>%
  #left_join(keys$bec, by = c("bec" = "zone")) %>%
  pivot_longer(-strata) %>%
  mutate(name = str_to_title(name)) %>%
  mutate(value = ifelse(value < 0, 0, value),
         class = ifelse(strata %in% forest_types, 
                        "Forest Types", 
                        "BEC Zones")) %>%
  ggplot(aes(x = value, y = name)) +
  stat_halfeye(
    # adjust bandwidth
    adjust = .5,
    # move to the right
    justification = -0.2,
    # remove the slub interval
    .width = 0,
    point_colour = NA
    ) +
  geom_boxplot(
    width = 0.12,
    # removing outliers
    outlier.color = NA,
    alpha = 0.5
  ) +
  stat_dots(
    # ploting on left side
    side = "left",
    # adjusting position
    justification = 1.1,
    # adjust grouping (binning) of observations
    binwidth = 0.0025,
    #position = "stack",
    # aes(fill = group, col = group),
    alpha = 0.5,
    aes(shape = class)
  ) +
  labs(y = NULL,
       x = "% Variance Explained",
       fill = "BEC Group",
       col = "BEC Group") +
  scale_x_continuous(labels = scales::label_percent()) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.position = c(0.8, .2))

@@

```{r path-setup}
path_base <- tibble(x1 = c(0,
             0,
             0,
             0,
             1),
       y1 = c(1,
              -1,
              1,
              -1,
              0),
       x2 = c(1.85,
              1.85,
              .85,
              .85,
              1.85),
       y2 = c(.15,
              -.15,
              .15,
              -.15,
              0)) %>%
  ggplot(aes(x = x1, y = y1)) +
  geom_segment(aes(xend = x2, yend = y2), arrow = arrow()) +
  geom_point(data = tibble(x1 = c(.85, .85), 
                           y1 = c(-.5, .5)),
             size = 60,
             shape = 21,
             col = "grey") +
  # geom_text(aes(x = 0, y = -1.15), label = "Modelled\nAttributes") +
  # geom_text(aes(x = 0, y = 1.15), label = "Extracted\nAttributes") +
  # geom_text(aes(x = .65, y = 0), label = "Overlap") +
  # geom_text(aes(x = 2, y = 0), label = "Dynamic\nHabitat\nIndices") +
  theme_void() +
  scale_x_continuous(limits = c(-0.1, 2.1))

ggsave(here::here("outputs", "path_base.png"), path_base)

@@

path_base +
  geom_text(data = path_r2_clean, aes(label = r2_label))

ggsave(here::here("outputs", "path_venn.png"),
        height = 3.5, width = 10)
```

@@

@@

      anova_out$`Pr(>F)` %>%
        as.data.frame() %>%
        rename(anova_p = 1) %>%
        drop_na() %>%
        mutate(anova_sig = cut(
          anova_p,
          breaks = c(0, 0.001, 0.01, 0.05, .10, 1),
          labels = c("***", "**", "*", ".", "NS")))  %>%
        mutate(axis = paste0("RDA", row_number())) %>%
        relocate(axis)
  }, .progress = "anovas"))

anova_sig_axes <- anova_outs %>%
  select(strata, anova) %>%
  unnest(cols = anova) # %>%
  # filter(anova_sig != "NS")

axis_varexp <- anova_sig_axes %>%
  inner_join(eig_outs)

axis_varexp %>%
  filter(anova_sig != "NS") %>%
  filter(strata %in% forest_types)

axis_varexp %>%
  filter(anova_sig != "NS") %>%
  mutate(category = ifelse(strata %in% forest_types, "Forest Types", "BEC Zones")) %>%
  # left_join(keys$bec, by = c("bec" = "zone")) %>%
  ggplot(aes(x = axis, y = var_exp, col = category, shape = category)) +
  geom_jitter(height = 0, width = 0.05) +
  theme(legend.position = "bottom") +
  labs(x = NULL,
       y = "% of Variance Explained",
       col = NULL,
       shape = NULL) +
  scale_y_continuous(labels = scales::label_percent())
# no secondary axis in alpine & coast.

axis_varexp %>%
  group_by(strata) %>%
  # filter(n() == 2) %>%
  select(strata, axis, var_exp) %>% 
  pivot_wider(names_from = axis,
              values_from = var_exp) %>%
  mutate(ratio = RDA1 / RDA2)

# smaller ratios present when the RDA1 doesn't explain much. when variance explained
# is low, the ratio is low. this is because there isn't much variance explained
```

```{r varexp-coniferous}
rda_varexp <- outs %>%
  mutate(varexp = map(rda, .f = \(rda) {
    rda_sum <- summary(rda)
    var_exp <- rda_sum$constr.chi / rda_sum$tot.chi
  })) %>%
  unnest(varexp) %>%
  arrange(desc(varexp)) %>%
  select(strata, varexp)

rda_varexp %>%
  arrange(strata %in% forest_types,
          desc(varexp))

source(here::here("scripts", "get_bec_proportion_vlce.R"))

prop_con <- bec_normal %>% 
  select(bec, category) %>% 
  group_by(bec) %>% 
  summarize(per_sample_for = mean(category == "Coniferous")) %>% 
  left_join(per_cons %>% 
              select(-value)) %>% 
  mutate(for_dif = per_sample_for - per_con) %>%
  arrange(desc(abs(for_dif))) %>%
  select(-for_dif) %>%
  pivot_longer(starts_with("per"))

con_var_plot <- prop_con %>%
  filter(name == "per_con") %>%
  left_join(rda_varexp, by = c("bec" = "strata")) %>% 
  left_join(keys$bec, by = c("bec" = "zone")) %>%
  mutate(name = ifelse(name == "per_con", 
                       "Global Percent Coniferous", 
                       "Sample Percent Coniferous")) %>%
  ggplot(aes(x = value, y = varexp)) +
  geom_point(aes(col = group)) +
  geom_smooth(method = "lm") +
  scale_x_continuous(labels = scales::label_percent()) +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(x = "Percent Coniferous",
       y = "Proportion of variance explained",
       col = NULL) +
  theme(legend.position = c(0.1, 0.85))

ggsave(here::here("outputs", "con_var.png"), plot = con_var_plot)
```

```{r bec-loadings}
out_loadings <- outs %>%
  mutate(loadings = map(rda, .f = function(rda) {
    summary(rda)$sites %>%
      as.data.frame() %>%
      select(starts_with("RDA"))
  }),
  loadings = map2(.x = loadings, .y = data, .f = bind_cols),
  loadings = map(loadings, .f = function(x) {
    rcorr <- x %>% select(-any_of(c("bec", "category"))) %>%
      as.matrix() %>%
      Hmisc::rcorr()

@@

  )) %>%
  select(strata, loadings) %>%
  unnest(loadings)
```

```{r}
sig_loads <- out_loadings %>% 
  mutate(loading_sig = cut(
          axis_p,
          breaks = c(0, 0.001, 0.01, 0.05, .10, 1),
          labels = c("***", "**", "*", ".", "NS"),
          right = F)) %>%
  inner_join(anova_sig_axes) %>%
  #filter(axis_sig == "NS") %>%
  arrange(desc(abs(axis_loading))) %>%
  filter(!(var %in% c("percentage_first_returns_above_mean", 
                    "elev_mean", 
                    "elev_stddev"))) %>%
  drop_na()# removes ima mindhi which has no variance

sig_loads %>%
  filter(anova_sig != "NS") %>%
  filter(endsWith(var, "DHI"),
         axis == "RDA2") %>%
  arrange(strata) %>%
  select(strata, var, axis_loading) %>%
  pivot_wider(names_from = var,
              values_from = axis_loading)
# when a second axis is present, it has a weak loading on the cumulative DHI, as compared to the other DHIs
# while the majoirty of the variability is contained within all three DHIs together, the second axis, when present, is more linked to the minmimum and variation DHI.

sig_loads %>%
  select(strata, var, axis, axis_loading, loading_sig) %>%
  filter(endsWith(var, "DHI"),
         loading_sig != "NS") %>%
  ggplot(aes(x = var, y = abs(axis_loading), group = strata)) +
  geom_point(aes(col = strata)) +
  geom_line(aes(col = strata), alpha = 0.5, lty = "dashed") +
  geom_hline(yintercept = 0, col = "red", alpha = 0.2) +
  facet_wrap(~axis) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  labs(x = NULL,
       y = "Axis Loading")

sig_loads %>%
  select(strata, var, axis, axis_loading) %>%
  filter(!endsWith(var, "DHI")) %>%
  left_join(keys$continuous, by = c("var" = "variable")) %>%
  ggplot(aes(x = var_long, y = abs(axis_loading), group = strata)) +
  geom_point(aes(col = strata)) +
  geom_line(aes(col = strata), alpha = 0.5, lty = "dashed") +
  geom_hline(yintercept = 0, col = "red", alpha = 0.2) +
  facet_wrap(~axis) +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(x = NULL,
       y = "Axis Loading (Absolute Value)")

sig_loads %>%
  filter(!endsWith(var, "DHI"),
         loading_sig != "NS") %>%
  group_by(strata, axis) %>%
  mutate(loading_rank = data.table::frankv(abs(axis_loading), 
                                           order = -1)) %>%
  ggplot(aes(x = fct_reorder(var, loading_rank),
             y = loading_rank, col = strata)) +
  geom_jitter(height = 0) +
  facet_wrap(~axis) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1)) +
  labs(x = NULL,
       y = "Axis Loading Rank")

# need to remove the correlations for variables that arent actually involved
# percentage_first_returns_above_mean, elev_mean, elev_stddev
out_loadings %>% count(var)

clean_loads <- sig_loads %>%
  filter(anova_sig != "NS",
         loading_sig != "NS") %>%
  bind_rows(all_data_loadings %>%
              mutate(strata = "All Data") %>%
              rename(loading_sig = sig)) %>%
  select(strata:axis_loading) %>%
  mutate(axis_loading = abs(axis_loading))
```

##### run to here for analysis to be done
radar chart based on groups, not hierarchical clustering

```{r radar}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

n = 8
cols = c4a(palette = "rainbow", n = 8)

radar_loads <- clean_loads %>%
  pivot_wider(names_from = var, 
              values_from = axis_loading) %>%
  left_join(keys$bec, by = c("strata" = "zone")) %>%
  mutate(group = ifelse(is.na(group), "Forest Types", group),
         group = ifelse(strata == "All Data", "All Data", group),
         group = ifelse(axis == "RDA2", "RDA2", group)) %>%
  group_by(group) %>%
  summarize(across(CumDHI:elev_p95, \(x) mean(x, na.rm = T))) %>%
  mutate(group = fct_relevel(group, "RDA2", after = Inf),
         group = fct_relevel(group, "Forest Types", after = 1))

radar_labels <- radar_loads %>%
  select(percentage_first_returns_above_2m:elev_p95) %>%
  colnames() %>%
  tibble(variable = .) %>%
  left_join(keys$continuous) %>%
  pull(var_long)

in_radar <- radar_loads %>%
  select(!ends_with("DHI")) %>%
  replace(is.na(.), 0) %>%
  ggradar(axis.labels = radar_labels,
          legend.title = "Cluster",
          values.radar = c("", "", 1),
          base.size = 1,
          background.circle.transparency = 0,
          group.point.size = 2,
          group.colours = cols,
          axis.label.size = 3,
          plot.extent.x.sf = 1.5,
          group.line.width = 1) +
  theme(legend.position = "none")

out_radar <- radar_loads %>%
  select(group, ends_with("DHI")) %>%
  replace(is.na(.), 0) %>%

  ggradar(legend.title = "Cluster",
          values.radar = c("", "", 1),
          base.size = 1,
          background.circle.transparency = 0,
          group.point.size = 2,
          group.colours = cols,
          axis.label.size = 3,
          plot.extent.x.sf = 1.5,
          group.line.width = 1) +
  theme(legend.position = "none")

boxplots <- clean_loads %>% 
  left_join(keys$continuous, by = c("var" = "variable")) %>%
  mutate(var_long = ifelse(is.na(var_long), var, var_long)) %>%
  select(!(divide_by:unit)) %>%
  left_join(keys$bec, by = c("strata" = "zone")) %>%
  mutate(group = ifelse(is.na(group), "Forest Types", group),
         group = ifelse(axis == "RDA2", "RDA2", group),
         group = ifelse(strata == "All Data", "All Data", group),
         inout = ifelse(endsWith(var, "DHI"), "Response", "Predictor")) %>%
  mutate(group = fct_relevel(group, "RDA2", after = Inf),
         group = fct_relevel(group, "Forest Types", after = 1)) %>%
  ggplot(aes(x = abs(axis_loading), y = fct_rev(fct_reorder(var_long, abs(axis_loading))))) +
  geom_boxplot(outlier.colour = "#00000000", width = .5) +
  geom_jitter(height = 0.1, aes(col = group, shape = group)) +
  facet_grid(cols = vars(axis),
             rows = vars(inout),
             scales = "free_y",
             space = "free_y") +
  labs(x = "Axis Loading (absolute)",
       y = NULL,
       col = NULL,
       shape = NULL) +
  theme(panel.spacing.y = unit(0, "lines"),
        legend.position = "right") +
  scale_shape_manual(values = c(17, 15, 16, 16, 16, 16, 16, 16)) +
  scale_colour_manual(values = cols)

radar_boxplot <- ((in_radar / out_radar) | boxplots)  +
  plot_annotation(tag_levels = "A") +
  plot_layout(guides = "collect") &
  theme(plot.tag = element_text(size = 12))

ggsave(here::here("outputs", "radar_boxplot.png"),
       plot = radar_boxplot,
       device = "png")
```

```{r drawio}

# make dir
drawio_dir <- here::here("outputs", "drawio")

dir.create(drawio_dir, showWarnings = F)

# templates
one_rda <-
  read_lines(here::here(drawio_dir, "_rda_path-text_replace-RDA1.drawio"))

two_rda <-
  read_lines(here::here(drawio_dir, "_rda_path-text_replace-RDA2.drawio"))

# list to iterate
stratum <- clean_loads %>%
  pull(strata) %>%
  unique()

# make sure everything gets change in rda fig
possibles <- clean_loads %>%
  select(var, axis) %>%
  complete() %>%
  mutate(possible = paste(var, axis, sep = "-")) %>%
  pull(possible) %>%
  unique()

# fractions dfs
fract_all <- path_r2 %>% 
  pivot_wider(names_from = type, values_from = Adj.R.squared) %>% 
  select(-residual) %>%
  mutate(strata = "All Data") %>% 
  relocate(strata)

fracts_all <- fracts_out %>%
  bind_rows(fract_all)

# variance explained by axis

all_varexp <- rda_eig %>%
  filter(name != "RDA3") %>%
  mutate(strata = "All Data") %>%
  rename(axis = name) %>%
  select(-value)

varexp_all <- axis_varexp %>%
  filter(anova_sig != "NS") %>%
  select(strata, axis, prop, var_exp) %>%
  bind_rows(all_varexp)

for (s in stratum) {
  strata_tp <- clean_loads %>%
    filter(strata == s) %>%
    mutate(textreplace = paste(var, axis, sep = "-"),
           axis_loading = as.character(round(axis_loading, digits = 2)))

  if (strata_tp %>% pull(axis) %>% unique() %>% length() == 1) {
    strata_drawio <- one_rda
  } else {
    strata_drawio <- two_rda
  }

  textreplace <- strata_tp %>%
    pull(textreplace)

  replaceval <- strata_tp %>%
    pull(axis_loading)

  for (i in 1:length(textreplace)) {
    text <- textreplace[[i]]
    val <- replaceval[[i]]

    strata_drawio <- str_replace(strata_drawio, text, val)
  }

  strata_drawio <- str_replace(strata_drawio, "title_strata", s)

  #strata_drawio <- strata_drawio[!str_detect(strata_drawio, "RDA")]
  # filter through all possibilities remaining, replace them with NS

  for(p in possibles) {
    strata_drawio <- str_replace(strata_drawio, p, "NS")
  }

  strata_fract <- fracts_all %>%
    filter(strata == s) %>%
    pivot_longer(extracted:overlap) %>%
    mutate(value = (value * 100) %>%
             round(1) %>%
             paste0(., "%"))

  fract_names <- strata_fract %>%
    pull(name)

  fract_value <- strata_fract %>%
    pull(value)

  for(i in 1:length(fract_names)) {
    strata_drawio <- str_replace(strata_drawio, fract_names[[i]], fract_value[[i]])
  }

  strata_varexp <- varexp_all %>%
    filter(strata == s) %>%
    mutate(var_exp = (var_exp * 100) %>%
             round(1) %>%
             paste0(., "%"))

  axis_names <- strata_varexp %>%
    pull(axis) %>%
    paste("var_exp", ., sep = "-")

  axis_vals <- strata_varexp %>%
    pull(var_exp)

  for(i in 1:length(axis_names)) {
    strata_drawio <- str_replace(strata_drawio, axis_names[[i]], axis_vals[[i]])
  }

  write_lines(strata_drawio, here::here(drawio_dir, paste0(s, ".drawio")))
}
```

```{r complexity-iqr}
strata_nested %>% 
  unnest(data) %>% 
  ggplot(aes(x = fct_reorder(strata, elev_cv, IQR), y = elev_cv)) + 
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# rgb map

```{r}
bec_dissolve <- read_sf(here::here("data", "shapefiles", "bec_dissolve.shp")) %>%
  janitor::clean_names() %>%
  select(zone)

# non normalized
loadings_rgb <- clean_loads %>% 
  group_by(axis) %>%
  pivot_wider(names_from = var,
              values_from = axis_loading) %>%
  ungroup() %>%
  replace(is.na(.), 0) %>%
  mutate(across(elev_cv:elev_p95, \(x) ifelse(x > 0.5, 0.5, x))) %>%
  mutate(dhi_rgb = rgb(CumDHI, VarDHI, MinDHI),
         pri_rgb = rgb(elev_p95, percentage_first_returns_above_2m, elev_cv, maxColorValue = 1),
         mod_rgb = rgb(basal_area, total_biomass, gross_stem_volume, maxColorValue = 1)) %>%
  select(strata, axis, percentage_first_returns_above_2m, elev_p95, elev_cv, ends_with("rgb")) %>%
  pivot_longer(ends_with("rgb"))

# normalized to max
loadings_rgb <- clean_loads %>% 
  group_by(var) %>%
  mutate(axis_loading = axis_loading / max(axis_loading)) %>%
  group_by(axis) %>%
  pivot_wider(names_from = var,
              values_from = axis_loading) %>%
  ungroup() %>%
  replace(is.na(.), 0) %>%
  #mutate(across(elev_cv:elev_p95, \(x) ifelse(x > 0.5, 0.5, x))) %>%
  mutate(dhi_rgb = rgb(CumDHI, VarDHI, MinDHI),
         pri_rgb = rgb(elev_p95, percentage_first_returns_above_2m, elev_cv, maxColorValue = 1),
         mod_rgb = rgb(basal_area, total_biomass, gross_stem_volume, maxColorValue = 1)) %>%
  #select(strata, axis, percentage_first_returns_above_2m, elev_p95, elev_cv, ends_with("rgb")) %>%
  pivot_longer(ends_with("rgb"))

becs <- bec_normal %>%
  pull(bec) %>%
  unique()
